# Excercise 4

## access the MASS package
```{r}
library(MASS)
```

## load the data
```{r}
data("Boston")
```

## explore the dataset
```{r}
str(Boston)
summary(Boston)
```



```{r}
library(tidyr)
library(corrplot)
library(dplyr)
```

## calculate the correlation matrix and round it
```{r}
cor_matrix<-cor(Boston) %>% round(digits = 2 )
```

## print the correlation matrix
```{r}
cor_matrix
```

## visualize the correlation matrix
```{r}
corrplot(cor_matrix, method="circle", type="upper", cl.pos="b", tl.pos="d", tl.cex = 0.6)
```

**It seems that "dis" is closely related to "age", "nox", "indus" and "zn".**

## center and standardize variables
```{r}
boston_scaled <- scale(Boston)
```

## summaries of the scaled variables
```{r}
summary(boston_scaled)
```

## class of the boston_scaled object
```{r}
class(boston_scaled)
```

## change the object to data frame
```{r}
boston_scaled <- as.data.frame(boston_scaled)
```

## summary of the scaled crime rate
```{r}
summary(boston_scaled$crim)
```

## create a quantile vector of crim and print it
```{r}
bins <- quantile(boston_scaled$crim)
```

```{r}
bins
```

## create a categorical variable 'crime'
```{r}
crime <- cut(boston_scaled$crim, breaks = bins, include.lowest = TRUE, labels = c("low", "med_low", "med_high", "high"))
```

## look at the table of the new factor crime
```{r}
table(crime)
```

## remove original crim from the dataset
```{r}
boston_scaled <- dplyr::select(boston_scaled, -crim)
```

## add the new categorical value to scaled data
```{r}
boston_scaled <- data.frame(boston_scaled, crime)
```

## number of rows in the Boston dataset 
```{r}
n <- nrow(boston_scaled)
```

## choose randomly 80% of the rows
```{r}
ind <- sample(n,  size = n * 0.8)
```

## create train set
```{r}
train <- boston_scaled[ind,]
```

## create test set 
```{r}
test <- boston_scaled[-ind,]
```

## save the correct classes from test data
```{r}
correct_classes <- test$crime
```

## remove the crime variable from test data
```{r}
test <- dplyr::select(test, -crime)
```

## linear discriminant analysis
```{r}
lda.fit <- lda(crime ~ ., data = train)
```

## print the lda.fit object
```{r}
lda.fit
```

## the function for lda biplot arrows
```{r}
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "orange", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}
```

## target classes as numeric
```{r}
classes <- as.numeric(train$crime)
```

## plot the lda results
```{r}
plot(lda.fit, dimen = 2, col = classes, pch = classes)
lda.arrows(lda.fit, myscale = 1)
```

## predict classes with test data
```{r}
lda.pred <- predict(lda.fit, newdata = test)
```

## cross tabulate the results
```{r}
table(correct = correct_classes, predicted = lda.pred$class)
```


## load MASS and Boston
```{r}
library(MASS)
data('Boston')
```

## euclidean distance matrix
```{r}
dist_eu <- dist(Boston)
```

## look at the summary of the distances
```{r}
summary(dist_eu)
```

## manhattan distance matrix
```{r}
dist_man <- dist(Boston, method = 'manhattan')
```

## look at the summary of the distances
```{r}
summary(dist_man)
```

## k-means clustering
```{r}
km <-kmeans(Boston, centers = 3)
```

## plot the Boston dataset with clusters
```{r}
pairs(Boston, col = km$cluster)
```
## Boston dataset is available
```{r}
set.seed(123)
```

## determine the number of clusters
```{r}
k_max <- 10
```

## calculate the total within sum of squares
```{r}
twcss <- sapply(1:k_max, function(k){kmeans(Boston, k)$tot.withinss})
```

## visualize the results
```{r}
library(ggplot2)
```

```{r}
qplot(x = 1:k_max, y = twcss, geom = 'line')
```

## k-means clustering
```{r}
km <-kmeans(Boston, centers = 2)
```

## plot the Boston dataset with clusters
```{r}
pairs(Boston, col = km$cluster)
```

## ggpairs
```{r}
library(GGally)
ggpairs(boston_scaled)
```

## Super bonus
```{r}
model_predictors <- dplyr::select(train, -crime)
```

### check the dimensions
```{r}
dim(model_predictors)
dim(lda.fit$scaling)
```

### matrix multiplication
```{r}
matrix_product <- as.matrix(model_predictors) %*% lda.fit$scaling
matrix_product <- as.data.frame(matrix_product)
```

```{r}
library(plotly)
```

### 3D figure
```{r}
plot_ly(x = matrix_product$LD1, y = matrix_product$LD2, z = matrix_product$LD3, type= 'scatter3d', mode='markers', color = train$crime )
```

