
# Excercise 6

## Read BPRSL data
```{r}
BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep=" ", header=TRUE)
str(BPRS)
```

```{r}
summary(BPRS)
```



**These data are treatment and groups**


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Convert the categorical variables of both sets to factors
```{r}
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
```
```{r}
str(BPRS)
```

## Convert to long form
```{r}
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(weeks = as.integer(substr(weeks, 5,5)))
```
## See the BPRS
```{r}
glimpse(BPRSL)
```


## Read RATS data
```{r}
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep="\t", header=TRUE)
str(RATS)
```
```{r}
RATS$Group <- factor(RATS$Group)
RATS$ID <- factor(RATS$ID)
```
```{r}
str(RATS)
```

## Convert to long form
```{r}
RATSL <-  RATS %>% gather(key = WDs, value = rats, -ID, -Group) %>%
  mutate(Time = as.integer(substr(WDs, 3, 4))) 
```
## See the RATS
```{r}
glimpse(RATSL)
```


```{r}
write.csv(RATS, "data/rats_long.csv")
write.csv(BPRS, "data/bprs_long.csv")
```

## Graphical Displays of Longitudinal Data



## Draw the plot
```{r}
ggplot(BPRSL, aes(x = weeks, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```




## Standardise the variable bprs
```{r}
BPRSL <- BPRSL %>%
  group_by(weeks) %>%
  mutate(stdbprs = (bprs - mean(bprs))/sd(bprs)) %>%
  ungroup()
```
## Plot again with the standardised bprs
```{r}
ggplot(BPRSL, aes(x = weeks, y = stdbprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")
```




## Summary Measure Analysis of Longitudinal Data

## Number of weeks, baseline (week 0) included

```{r}
n <- BPRSL$weeks %>% unique() %>% length()
```
## Summary data with mean and standard error of bprs by treatment and week 
```{r}
BPRSS <- BPRSL %>%
  group_by(treatment, weeks) %>%
  summarise( mean = mean(bprs), se = sd(bprs)/sqrt(n) ) %>%
  ungroup()
```

## Glimpse the data
```{r}
glimpse(BPRSS)
```


Mean response profiles for the two treatment groups in the BPRS data.




## Plot the mean profiles
```{r}
ggplot(BPRSS, aes(x = weeks, y = mean, linetype = treatment, shape = treatment)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(bprs) +/- se(bprs)")
```

```{r}
RATSL<-gather(RATS, key=WDs, value=weight, WD1:WD64)%>%mutate(time= as.integer(substr(WDs,3,4)))
```

```{r}
ggplot(RATSL, aes(x = time, y = weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$weight), max(RATSL$weight)))
```
Standardise the weight
```{r}
RATSL<- RATSL%>%group_by(time)%>%mutate(stdweight = (weight - mean(weight))/sd(weight))%>%ungroup()
glimpse(RATSL)
```
```{r}
p1<-ggplot(RATSL, aes(x=time, y=stdweight, linetype=ID))
p2<-p1+geom_line()+scale_linetype_manual(values=rep(1:10, times=4))
p3<-p2+facet_grid(. ~Group, labeller = label_both)
p4<-p3+theme_bw()+theme(legend.position ="none")
p5<-p4+theme(panel.grid.minor.y=element_blank())
p6<-p5+scale_y_continuous(name="standardized weights")
p6
```

Numberof times, baseline (time 0)
```{r}
n<-RATSL$time%>%unique()%>%length()
```
Make a summary data
```{r}
RATSS<-RATSL%>%group_by(Group, time)%>% summarise(mean=mean(weight), se=sd(weight)/sqrt(n))%>%ungroup()
glimpse(RATSS)
```
Visualize the  mean weight and SE
```{r}
p1<-ggplot(RATSS, aes(x=time, y=mean, linetype=Group, shape=Group))
p2<-p1+geom_line()+scale_linetype_manual(values=c(1,2,3))
p3<-p2+geom_point(size=3)+scale_shape_manual(values=c(1,2,3))
p4<-p3+geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3)
p5<-p4+theme_bw()+theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
p6<-p5+theme(legend.position= "top")
p7<-p6+scale_y_continuous(name="mean(weights) +/- se(weights)")
p7
```
