
# Excercise 5

## Read BPRSL data
```{r}
BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep=" ", header=TRUE)
str(BPRS)
```

```{r}
summary(BPRS)
```



**These data are treatment and groups**


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Convert the categorical variables of both sets to factors
```{r}
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
```
```{r}
str(BPRS)
```

## Convert to long form
```{r}
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(weeks = as.integer(substr(weeks, 5,5)))
```
## See the BPRS
```{r}
glimpse(BPRSL)
```


## Read RATS data
```{r}
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep="\t", header=TRUE)
str(RATS)
```
```{r}
RATS$Group <- factor(RATS$Group)
RATS$ID <- factor(RATS$ID)
```
```{r}
str(RATS)
```

## Convert to long form
```{r}
RATSL <-  RATS %>% gather(key = WDs, value = rats, -ID, -Group) %>%
  mutate(Time = as.integer(substr(WDs, 3, 4))) 
```
## See the RATS
```{r}
glimpse(RATSL)
```


```{r}
write.csv(RATS, "data/rats_long.csv")
write.csv(BPRS, "data/bprs_long.csv")
```

## Graphical Displays of Longitudinal Data



## Draw the plot
```{r}
ggplot(BPRSL, aes(x = weeks, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```




## Standardise the variable bprs
```{r}
BPRSL <- BPRSL %>%
  group_by(weeks) %>%
  mutate(stdbprs = (bprs - mean(bprs))/sd(bprs)) %>%
  ungroup()
```
## Plot again with the standardised bprs
```{r}
ggplot(BPRSL, aes(x = weeks, y = stdbprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")
```




## Summary Measure Analysis of Longitudinal Data

## Number of weeks, baseline (week 0) included

```{r}
n <- BPRSL$weeks %>% unique() %>% length()
```
## Summary data with mean and standard error of bprs by treatment and week 
```{r}
BPRSS <- BPRSL %>%
  group_by(treatment, weeks) %>%
  summarise( mean = mean(bprs), se = sd(bprs)/sqrt(n) ) %>%
  ungroup()
```

## Glimpse the data
```{r}
glimpse(BPRSS)
```


Mean response profiles for the two treatment groups in the BPRS data.

## Plot the mean profiles
```{r}
ggplot(BPRSS, aes(x = weeks, y = mean, linetype = treatment, shape = treatment)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(bprs) +/- se(bprs)")
```

