
# Excercise 2

# to see the data
```{r}
df <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", sep=",", header=TRUE)
```
# to see structure and dimmension
```{r}
str(df)
dim(df)
```
# to acess the ggplot 2 library
```{r}
library(ggplot2)
ggplot(df)
pairs(df[-1])
```
# to access the GGally and ggplot2
```{r}
library(GGally)
library(ggplot2)
```
# to see summary
```{r}
summary(df)
```
# create a more advanced plot matrix with ggpairs
```{r}
p <- ggpairs(df, maaping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))
```
# to draw the plot
```{r}
p
```
#*The three explanatory variables are "attitude", "strategy" and "surface".* **These are choosen based on r value**

# to see scatter plot
```{r}
qplot(attitude, points, data = df) + geom_smooth(method = "lm")
```
# fit a linear model
```{r}
model <- lm(points ~1, data = df)
```
# summary of the model
```{r}
summary(model)
```
# to create an plot matrix
```{r}
library(GGally)
ggpairs(df, lower = list(combo = wrap("facethist", bins = 20)))
```

# to create a regression model with multibple explanatory variables
```{r}
model2<- lm(points ~ attitude + stra + surf, data = df)
summary(model2)
```
**remove the surf because its p value was the highest**
```{r}
model3<- lm(points ~attitude + stra, data=df)
summary(model3)
```
**the most significant variable was attitude. Stra had only tendency**

# to draw diagnostic plots
```{r}
par(mfrow = c(2,2))
plot(model3, which = c(1,2,5))
```
**QQ plot showed this model is normally distributed**
**all looks good**
